---
title: "Week 6 Coding Lesson - Intro to Linear Regression"
author: "Michelle Shteyn"
date: "11/9/2020"
output: html_document
---

## Part 1: Attach Packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Attach packages
library(tidyverse)
library(palmerpenguins)
library(ggpubr)
library(broom)
```


## Part 2: A rank-based test example (Mann Whitney U)
```{r}
set.seed(1414)
gp_1 <- sample.int(20, size = 15, replace = TRUE)

set.seed(1424)
gp_2<-sample.int(30,size=15,replace = TRUE)
```

Look at each sample using hist().
```{r}
hist(gp_1)
hist(gp_2)
```

Perform Mann-Whitney U to answer is there a significant difference in ranks (medians) between gp_1 and gp_2? Using the wilcox.test() function. 
```{r}
my_mwu <- wilcox.test(gp_1, gp_2)
```


## Part 3: Simple linear regression

### A. Look at it!

This is always the FIRST thing we do.

Make exploratory scatterplot of penguin flipper length vs. body mass

```{r}
ggplot(data = penguins, aes(x=flipper_length_mm, y = body_mass_g))+
  geom_point()

```

## B. Model it

```{r}
#Linear model, stored as penguin_lm:
  penguin_lm <-lm(body_mass_g ~ flipper_length_mm, data=penguins)
  
# Return the complete overview:
summary(penguin_lm)
  
  
```
75.9 % of body mass is explained by flipper length. 

### C. Access model outputs
Trying to get all of the statistical information from the summary() function would be kind of a mess. 

We can use the broom: :tidy() function to get the model outputs in nice data frame format:

```{r}
penguin_lm_tidy<-broom::tidy(penguin_lm)
```

Produced a nice table of model outputs, which we can then refer to later on. Examples:
```{r}
#Get the intercept: 
penguin_int<-penguin_lm_tidy$estimate[1]
penguin_int

#Then to get the flipper_length coefficient:
penguin_coef<-penguin_lm_tidy$estimate[2]
penguin_coef

```
What about getting some other model information (df, F-stat, p-value, etc.)?
```{r}
# Metrics at a glance:
penguin_lm_out<-broom::glance(penguin_lm)
penguin_lm_out
```

We can use the results of both to write a statement about the model that will **automatically update** if anything about the model changes! Make sure to look at the .Rmd (not just this knitted html) to learn how to reference the outputs automatically in text. For example: 

"Simple linear regression was used to explore the relationship between penguin flipper length (mm) and body mass (g) across all three penguin species, and including both male and female penguins. A significant regression model was found ($\beta$ = `r round(penguin_coef,3)`, F(`r penguin_lm_out$df`,`r penguin_lm_out$df.residual`) = `r round(penguin_lm_out$statistic,1)`, p < 0.001) with an R^2^ of `r round(penguin_lm_out$r.squared,3)`."

**Note:** This might seem *really* tedious to write out, but the advantages are worth it. All values will be automatically updated when the model is updated! Reproducible and way less opportunity for human error. Plus, once you have this template statement made, you can reuse it for future regression models and just replace `penguin_lm_out` and `penguin_coef` with the appropriate objects for your new model! 

Note that I use "p < 0.001" here if the p-value is very small - this is somewhat standard. 

### D. Explore model assumptions

Use the plot() function on the model, which will automatically create four useful visualizations to consider assumptions!
```{r}
plot(penguin_lm)
```

## E. Visualize the model
```{r}
ggplot(data = penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point(size = 2) +
  geom_smooth(method = "lm",
              color = "red",
              size = 0.5,
              fill = "gray10",
              alpha = 0.5) +
  theme_light() +
  ggpubr::stat_regline_equation(label.x = 180, label.y = 5700)
```

### F. Find Pearson's *r* for correlation:

```{r}
penguins_cor <- cor.test(penguins$flipper_length_mm, penguins$body_mass_g)
```

Here, we see that there is a strong positive correlation between penguin flipper length and body mass (*r* = `r round(penguins_cor$estimate,2)`, t(`r penguins_cor$parameter`) = `r round(penguins_cor$statistic,2)`, p < 0.001). 

**Note**: Once you have a "template" statement, you can just replace `penguins_cor` here with whatever your correlation analysis is stored as! You don't need to recreate the wheel every time! 

## END LAB - CONGRATULATIONS!























